---
title: "Filters and Joins"
author: "James Bourke"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

library(tidyverse)
library(readxl)
library(here)
library(kableExtra)

```

Get some fish.csv and kelp_fronds.xlsx

```{r}
fish <- read_csv(here("data","fish.csv"))

kelp_abur <- read_excel(here("data","kelp_fronds.xlsx"),
                        sheet = "abur")

# default in `read_excel()` is to read in the first worksheet
# to read a different sheet, move outside of here(), and add an argument sheet = ""
```


### Filter by an exact match

Exact match: `==`

Example 1: only keep observations from `fish` where the common_name variable matches (==) "garibaldi"

```{r}
fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")
```

>, <, >=, <=, ==

Create a subset of fish called fish_over50, starting from fish, where the total_count is greater than OR equal to 50

```{r}
fish_over50 <- fish %>% 
  filter(total_count >= 50)
```


### What if we want to filter by this OR this?

You can use the vertical line `|` "or"

```{r}
fish_2sp <- fish %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith")

# If you're filtering for multiple matches within the same variable, you can use the %in%

fish_3sp <- fish %>% 
  filter(common_name %in% c("garibaldi", "blacksmith","black surfperch"))
```

Activity: subset from fish (fish_gar_2016) that keeps all observations where the year is 2016 OR the common_name is "garibaldi"

```{r}
fish_gar_2016 <- fish %>% 
  filter (year == 2016 | common_name == "garibaldi")
```

### Filter to match this AND that

, or & or piped filter statements

```{r}
aque_2018 <- fish %>% 
  filter(year == 2018, site == "aque")
```

To exclude things: `!=`

### stringr::str_detect()

```{r}
fish_black <- fish %>% 
  filter(str_detect(common_name, pattern = "black"))
```

Subset of data if the common name must include the string "it"?

```{r}
fish_it <- fish %>% 
  filter(str_detect(common_name, pattern = "it"))
```

### `dplyr::*join`

full_join() will keep everything across different data frames, even if there is not a complete match

```{r}
ab_kelpfish <- kelp_abur %>% 
  full_join(fish, by = c("year", "site"))
```

### `left_join()` to keep all from x, and merge matching things from y with it (exclude things from y that don't have a match in x)

```{r}
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year","site"))
```

### inner_join() to merge data frames, only keeping observations with a match in both

```{r}
kelp_fish_inner <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
```

### Pipe together multiple steps

- create an object called `my_fish_join`
- starting from `fish`
- filtering to only include observations at site "abur" from the year 2017
- join the kelp_abur data frame to the resulting subset using left_join()
- add a new column (`mutate()`) to calculate fish per kelp frond density (counts of fish / total fronds)

```{r}
my_fish_join <- fish %>% 
  filter(year == 2017, site == "abur") %>% 
  left_join(kelp_abur, by = c("year","site")) %>% 
  mutate(fish_per_frond = total_count/total_fronds)
  
```

Nice HTML table using kable + kableExtra

```{r}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```

### Create column graph with common_name on x and fish_per_frond on y

```{r}
ggplot(data = my_fish_join, aes(x = common_name, y = fish_per_frond)) +
  geom_col(
    aes(fill = common_name))
```

