---
title: "Creating graphs with ggplot2"
author: "James Bourke"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.

We will install these packages:

- tidyverse: suite of packages for wrangling and visualization (ggplot)
- readxl: to read in excel files (in many ways)
- here: to help with filepaths

```{r setup, message=FALSE}
library(tidyverse) #install.packages("tidyverse")
library(readxl) #install.packages("readxl")
library(here) #install.packages("here")
```

# California National Parks Data

```{r read-data}

ca_np <- read_csv(here("data/ca_np.csv"))
# also be written: ca_np <- read_csv("data/ca_np.csv)

ci_np <- read_excel(here("data/ci_np.xlsx"))

```

### Make my own first graph!

We're going to make a line plot of year (x-variable) vs. visitors (y-variable) in ggplot:

```{r line-graph, message=FALSE,warning=FALSE,echo=FALSE}

ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

# update the geom to geom_point(), geom_col(), geom_area()

```

Make a base layer for my CI plots:

```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point()
```

### Now let's start customizing!

- Line / point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha

```{r customize,echo=false}

gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )

```

Activity: create a ggplot scatterplot (geom_point) that has triangle symbols (hint: ?pch), of a non-default size, and that are an updated color

```{r}

gg_base +
  geom_point(
    pch = 2,
    size = 4,
    color = "green"
  )

```

### Update graph aesthetics based on variables

```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = year),
    pch = 5
  )
```

### Use gg-themes for overall updates

- theme_light()
- theme_minimal()
- theme_bw()

```{r}
gg_base +
  geom_area() +
  theme_light()
    
```

### Create a multi-series line plot using data for all CA National Parks

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line(
    aes(color = park_name)
  )
```

```{r}
my_np_graph <- ggplot(data = ca_np, aes(x = year, y = visitors)) +
  geom_point(aes(color = park_name), show.legend = FALSE) +
  facet_wrap(~park_name, scales = "free")
```

Export my graph as a .png
```{r}
ggsave(plot = my_np_graph, here("figures", "np_graph.png"))
```

